Customer holdrisk sampling distribution based on line probabilities
========================================================
    
    ```{r, message = F}
#load packages
set.seed(123)
require(ggplot2)
require(reshape)
require(scales)
```

Compare 3 bettors who all have 3% ROI, first customer bets only 10% dogs,
second customer bets 50% coinflips and last customer 90% favorites.
Each customer has sample size of 1000 bets

ROI = return on investment, defined as winloss/wagered. Hold is defined as winloss/volume, where volume is min(wagered,towin). So betting favorites volume is always lower than wagered. And as a result hold is amplified compared to holdrisk for big favorites. In the querybuilder ROI is called holdrisk.

First we'll simulate some games
```{r}
# Real probability of lines for the simulation
lineprob <- c("10% dog" = 0.1, "coinflip" = 0.5, "90% fav" = 0.9)

# Customer ROI, (same thing as holdrisk), 0.03 means 3%, defined as Winloss/RiskAmount (instead of Volume)
ROI <- 0.03
# Assume customer wagers 1 dollar per line, so samplesize also equals to total amount wagered
# Note that customer is risking 1 unit, so when betting favorites the volume is smaller
samplesize <- 2000
number_of_simulations <- 50000
# Line odds in decimal odds
decimal_odds <- (1+ROI)/lineprob

# simulate the game results (won/loss)
flips <- sapply(lineprob, function(x) rbinom(number_of_simulations, samplesize, x))
```

Now bookie sees these results and tries to figure out the real probabilities of the games and customers holdrisk/hold

# Game probabilities

```{r}
# sampling distribution for the mean of real probabilities
simulatedlineprob <- flips/samplesize

# length of the confidence intervals for the probabilities of the game
apply(simulatedlineprob, 2, function(x) quantile(x, 0.975) - quantile(x, 0.025))

```
In terms of absolute probabilities, the confidence interval for the coinflip
game probabilities is the widest. The length for the 10% dog and 90% fav game
confidence intervals has to be equal because of symmetry


```{r}
# Mean and confidence interval end points for the probabilities of the game
cbind("2.5%" = c(apply(simulatedlineprob, 2, function (x) quantile(x, 0.025))), "mean" = apply(simulatedlineprob, 2, mean), "97.5%" = c(apply(simulatedlineprob, 2, function (x) quantile(x, 0.975))))

melted_simulatedlineprob <- melt(simulatedlineprob)
colnames(melted_simulatedlineprob)[2] <- "Customer"
ggplot(melted_simulatedlineprob, aes(value, fill = Customer)) + geom_density(alpha = 0.2) +
ggtitle("Histogram of maximum likehood estimates for win probabilities") + xlab("Win probability") +
scale_x_continuous(label = percent)   
```

# Customer profits (winloss)

```{r}
# customers total profits = number of bets won * decimal odds - total amount wagered
simulatedwinloss <- t(t(flips)*decimal_odds) - samplesize
#apply(simulatedwinloss, 2, sd)

# length of the confidence intervals for winloss amount
apply(simulatedwinloss, 2, function(x) quantile(x, 0.975) - quantile(x, 0.025))
```




```{r}
# Confidence intervals for winloss amount
cbind("2.5%" = c(apply(simulatedwinloss, 2, function (x) quantile(x, 0.025))), "mean" = apply(simulatedwinloss, 2, mean), "97.5%" = c(apply(simulatedwinloss, 2, function (x) quantile(x, 0.975))))

melted_simulatedwinloss <- melt(simulatedwinloss)
colnames(melted_simulatedwinloss)[2] <- "Customer"
ggplot(melted_simulatedwinloss, aes(value, fill = Customer)) + geom_density(alpha = 0.2) +
ggtitle("Histogram of customer winloss amounts") + xlab("Customer winloss") +
scale_x_continuous(label = dollar)   




# holdrisk is just winloss/total amount wagered
simulatedholdrisk <- simulatedwinloss/(samplesize)
#apply(simulatedholdrisk, 2, mean)
#apply(simulatedholdrisk, 2, sd)

# length of the confidence intervals for holdrisk
apply(simulatedholdrisk, 2, function(x) quantile(x, 0.975) - quantile(x, 0.025))


# Mean and confidence interval for holdrisk
cbind("2.5%" = c(apply(simulatedholdrisk, 2, function (x) quantile(x, 0.025))), "mean" = apply(simulatedholdrisk, 2, mean), "97.5%" = c(apply(simulatedholdrisk, 2, function (x) quantile(x, 0.975))))

melted_simulatedholdrisk <- melt(simulatedholdrisk)
colnames(melted_simulatedholdrisk)[2] <- "Customer"
ggplot(melted_simulatedholdrisk, aes(value, fill = Customer)) + geom_density(alpha = 0.2) +
ggtitle("Histogram of calculated hold risks") + xlab("Customer hold risk percentage") +
scale_x_continuous(label = percent)   



# Note that the confidence interval for 10% dog game probabilities was narrower than
# the interval for coinflip games. But here when looking at the customer holdrisk's
# intervals, the interval for 10% dog bettor is wider than for the coinflip bettor

# Comparing the 10% dog and 90% favorite bettor, even tho we had same confidence
# interval for the game probabilities, the hold risk probabilities for these two
# customers are totally different

# These are explained by the fact that customer's holdrisk is calculated by
# ROI = 1 - decimal_odds * lineprob. When betting 10% dogs the decimal_odds are
# roughly 10, so even a small change in lineprobability will change the ROI
# pretty big, because we're multiplying the line probability by decimal_odds.
# Compare this to the guy betting 90% favorites, his decimal_odds
# are ~ 1.11, so in fact his ROI is capped by 11% even if he was betting
# fixed games.

# hold
simulatedhold <- simulatedwinloss/(samplesize)
# for hold, we have to multiply the 90% fav holdrisks by 10, because risk is 1 but volume only 0.1
simulatedhold[,3] <- simulatedhold[,3]*10
#apply(simulatedhold, 2, mean)
#apply(simulatedhold, 2, sd)

# Mean and confidence interval for hold, note that
cbind("2.5%" = c(apply(simulatedhold, 2, function (x) quantile(x, 0.025))), "mean" = apply(simulatedhold, 2, mean), "97.5%" = c(apply(simulatedhold, 2, function (x) quantile(x, 0.975))))

melted_simulatedhold <- melt(simulatedhold)
colnames(melted_simulatedhold)[2] <- "Customer"
ggplot(melted_simulatedhold, aes(value, fill = Customer)) + geom_density(alpha = 0.2) +
    ggtitle("Histogram of calculated holds") + xlab("Customer hold percentage") +
    scale_x_continuous(label = percent)   




# distance between 95% quantile and 5% quantile is the width of our confidence interval

# Because decimal_odds = ROI / lineprob, this means that  ROI = decimal_odds * lineprob.
# Take derivatives with respect lineprob -> delta ROI = decimal_odds.
# Basically the bigger decimal_odds are, the more leverage in his ROI customer gets when
# lineprobability changes.
```